---
title: "Tidy Transformations"
author: "Allison O'Connor"
date: "September 27, 2014"
output: html_document
---

This is a document to demonstrate how to transform messy data into Tidy Data.

First we load the libraries we need. 
```{r}
library(historydata)
library(tidyr)
library(dplyr)
library(lubridate)
```


Then we set the working directory and write the first transformation using the gather function.

```{r}

setwd("~/Documents/Clio3/rproject")
sarna <- read.csv("historydata/data-raw/sarna.csv", stringsAsFactors = FALSE)
sarna %>% 
    gather(type, numb, -year)
   
```

Here is cleaner version of the catholic.dioceses.geocoded.csv file.

```{r}
setwd("~/Documents/Clio3/rproject")
catholic_dioceses <- read.csv("historydata/data-raw/catholic.dioceses.geocoded.csv", stringsAsFactors = FALSE)
catholic_dioceses %>%
  select(diocese, 
         erected = date.erected,
         metropolitan = date.metropolitan,
         rite,
         lat = geo.lat,
         long = geo.lon) %>%
  gather (event, date, -diocese, -rite, -lat, -long) %>%
  tbl_df()

```



If we want to display only certain data, we can use the filter function.

```{r}
sarna <- read.csv("historydata/data-raw/sarna.csv", stringsAsFactors = FALSE)
sarna %>%
  gather(type, numb, -year) %>%
  filter(type == "estimate_low")

```



Here we take population data, group it by year, then add up the population for all states by year and display the number of states in that year.

```{r}
pop <- read.csv("historydata/data-raw/nhgis0011_ts_state.csv", stringsAsFactors = FALSE)
pop %>%
  group_by(YEAR) %>%
  summarize(total_pop = sum(A00AA, na.rm = TRUE),
            number_states = n())

```

```{r}


```
